Homicide Increases in the COVID-19 Era

### 2025-02-22 Draft1.qmd
Added TOC

fontsize: 10pt

geometry: margin=1in
bibliography: references.bib
---

@@

```{r}
# Primary datasets: Crime Data 2020-Present and 2010-2019
crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")

@@

# Plot 1 - Examining Crime Density from 2010–2019 and 2020-Present

@@

```
Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.

# Plot 2 - Examining Number Of Crimes Over Time

@@

# Using Regression to Predict ____

```
# Interpretation

# Plot 3 Crime rate frequency by day of the week

@@

# Plot bar chart with horizontal labels
barplot(crime_counts, col = "lightblue",
        main = "Crime Occurrences by Day of the Week",
        xlab = "Day of the Week", ylab = "Number of Crimes",
        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
```
# Interpretation

@@

# Plot 4 Gender Distribution for crimes over time

@@

# Create a pie chart using base R
color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
pie(
  gender_counts, 
  labels = paste(names(gender_counts), 
                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
  col = color[names(gender_counts)],
  main = "Distribution of Crime Victims by Gender"
)
```
# Interpretation

@@

# Plot 5 - Median Household Income by Zip Code in Los Angeles by Year

```{r}
library(sf)
library(tidygeocoder)
library(tidyverse)
library(RColorBrewer)

zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

median_household_income$S1903_C03_001E <- as.integer(median_household_income$S1903_C03_001E)

@@

@@

par(mfrow = c(2, 3), mar = c(2, 2, 2, 6), oma = c(0, 0, 4, 0))

@@

  # Plot
  plot(st_geometry(subset_data),
       col = palette[bins],
       border = 'grey',
       axes = FALSE,
       main = yr,
       cex.main = 1.5)

  # Add legend
  legend("right", inset = c(-0.5, 0), legend = round(quantile(subset_data$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0),
         fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
}
mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
```

# Plot 6 - Frequency of Top 10 Crimes

@@

>>>>

fontsize: 10pt
toc: true                   # add table of contents at the beginning
toc-depth: 3               # Only titles that start with # or ##
geometry: margin=1in
bibliography: references.bib
---

@@

```{r, eval=FALSE}
library(sf)
library(tidygeocoder)
library(tidyverse) # using only for joining datasets to build map visual
library(RColorBrewer)

# Primary datasets: Crime Data 2020-Present and 2010-2019
crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")

@@

# Plots

## Examining Crime Density from 2010–2019 and 2020-Present

@@

```

Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.

## Examining Number Of Crimes Over Time

@@

## Crime rate frequency by day of the week

@@

# Plot bar chart with horizontal labels
barplot(crime_counts, col = "lightblue",
        main = "Crime Occurrences by Day of the Week",
        xlab = "Day of the Week", ylab = "Number of Crimes",
        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
```

### Interpretation

@@

## Gender Distribution for crimes over time

@@

# Create a pie chart using base R
color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
pie(
  gender_counts, 
  labels = paste(names(gender_counts), 
                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
  col = color[names(gender_counts)],
  main = "Distribution of Crime Victims by Gender"
)
```

### Interpretation

@@

## Median Household Income by Zip Code in Los Angeles by Year

```{r}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

@@

zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)

@@

# setting a layout for a 2x3 grid, margins adjusted to fit the plots, oma() for outer margin area.
par(mfrow = c(2, 3), mar = c(1, 1, 1, 6), oma = c(0, 0, 3, 0), cex.main = 1.5)

@@

  # Plot
  plot(st_geometry(subset_data),
       col = palette[bins],
       border = 'grey',
       axes = FALSE)
  title(yr, line = -1)

  # Add legend
  legend("right", inset = c(-0.5, 0), legend = round(quantile(subset_data$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0),
         fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
}
mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1.5, font = 2)
```

## Frequency of Top 10 Crimes

@@

## Using Regression to Examine Weekly Crime Rates

```

### Interpretation
